Seshadri constants of curve configurations on surfaces

### 2024-03-20 sc-curve-config.tex
Generalised version-1-20th March (12:08)

	\section{Seshadri constants for certain transversal curve arrangements}\label{main}	
	%\subsection{General results for any surface}
	We work with curve configurations on surfaces satisfying the following assumption. 
	\begin{assumption}\label{star1}
		Let $\mathcal{C}=\{C_1,C_2, \ldots ,C_d\}$, $d\geq 4$ be a (connected) 
  transversal arrangement of curves 
		on a complex smooth projective surface $X$ such that all the curves 
  $C_i$ in $\mathcal{C}$ are linearly equivalent to a
			fixed divisor $A$ on $X$.
 % with the following properties: 
%		\begin{itemize}
%			\item There is no point where all the curves meet, i.e., $t_d=0.$
%			\item  All the curves $C_i$ in $\mathcal{C}$ are linearly equivalent to a
%			fixed divisor $A$ on $X$.
%		\end{itemize}
	\end{assumption}

@@

	For star arrangements, the base constant is exactly equal to $C_1^2(d-1)$.
	We now show that the lower bound in Question \ref{QuestM} can be achieved by certain star configuration of curves on $X.$	So Question \ref{question} has an affirmative answer for these star arrangements.

	\textcolor{red}{Let us denote $b_i =  \# \, C_{i} \cap {\rm Sing}(\mathcal{C})$. Note that since our arrangement is connected, $b_i >0$.}
	\begin{theorem}
		\label{prop:star} Let $X$ be a nonsingular projective surface and let $\mathcal{C} = \{C_{1}, \ldots, C_{d}\}$ be a star configuration of curves on $X$ with $d \geq 4$. 
		Let $L$ be an ample line bundle on $X$
		such that  $\frac{d\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L-C_i$ is nef for each $i$.   Then 
		%$$\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \inf_{C_i}{\frac{L\cdot C_i}{C_i^2(d-1)}}.$$
		$$\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \inf_{C_i}{\{\frac{L\cdot C_i}{b_i}\}}.$$
	\end{theorem}
	\begin{proof}
	 Let $\mathcal{P} = \{p_{1}, \ldots, p_{s}\}$ be the singular locus of  $\mathcal{C}$. Note that $p_{1}, \ldots, p_{s}$ are all double points of 
	  $\mathcal{C}$. Exactly $b_i$ of these points lie on $C_i$. So the Seshadri ratio computed by $C_i$ is precisely 
	  $ \frac{L\cdot C_i}{b_i}$. %Note that the same ratio is also contributed by each of the curves $C_2,\ldots, C_d$.

	\begin{theorem}
		\label{prop:star} Let $X$ be a nonsingular projective surface and let $\mathcal{C} = \{C_{1}, \ldots, C_{d}\}$ be a star configuration of curves on $X$ with $d \geq 4$ such that each $ C_i$ is linearly equivalent to a fixed divisor $A$ on $X$. Let $L$ be an ample line bundle on $X$
		such that  $\frac{dC_1^2}{(d-1)(L\cdot C_1)}L-C_1$ is nef.   Then 
		$$\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \frac{L\cdot C_1}{C_1^2(d-1)}.$$
	\end{theorem}
	\begin{proof}
	 Let $\mathcal{P} = \{p_{1}, \ldots, p_{s}\}$ be the singular locus of  $\mathcal{C}$. Note that $p_{1}, \ldots, p_{s}$ are all double points of 
	  $\mathcal{C}$. Exactly $C_1^2(d-1)$ of these points lie on $C_1$. So the Seshadri ratio computed by $C_1$ is precisely 
	  $ \frac{L\cdot C_1}{C_1^2(d-1)}$. Note that the same ratio is also contributed by each of the curves $C_2,\ldots, C_d$.

	 Suppose that there exists an irreducible and reduced curve $D$, different from $C_{i}$ for $i \in \{1,\ldots, d\}$, having multiplicity $m_{p}(D)$ at each point $p \in \mathcal{P}$ such that
		$$\frac{L \cdot D}{\sum_{p \in \mathcal{P}} m_{p}(D)} < \frac{L\cdot C_1}{C_1^2(d-1)}.$$
This gives
		$$(\triangle) : \quad (L \cdot D)C_1^2(d-1) < \left(\sum_{p \in \mathcal{P}} m_{p}(D)\right)L\cdot C_1.$$
		By B\'ezout's theorem, we obtain
		\begin{eqnarray*}
		D\cdot (C_{1}+ \ldots + C_{d})=d(D\cdot C_1) &\geq& \sum_{p \in \mathcal{P}}m_{p}(D)\cdot m_{p}(C_{1}+ \ldots + C_{d})\\
		&\stackrel{(*)}{\geq}& 2\sum_{p \in \mathcal{P}} m_{p}(D) \\
		&\stackrel{(\triangle)}{>}& \frac{2(L \cdot D)C_1^2(d-1)}{L\cdot C_1},
		\end{eqnarray*}
		where $(*)$ comes from the fact that all the singular points of $C_{1}+ \ldots + C_{d}$ are at least double points. Now we use the nefness 
		of $\frac{dC_1^2}{(d-1)(L\cdot C_1)}L-C_1$ and note the following  inequality: 
		\[
		\frac{dC_1^2}{(d-1)(L\cdot C_1)}L\cdot D \geq D\cdot C_1.
		\] 
		This leads to the following:
		\begin{eqnarray*}
			\frac{d^2C_1^2}{(d-1)(L\cdot C_1)}L\cdot D &\geq& d (D\cdot C_1) > \frac{2(d-1)(L\cdot D)C_1^2}{L\cdot C_1} \\
			\Rightarrow \frac{d^2C_1^2}{d-1} &>& 2(d-1)C_1^2 \quad \quad \text{(since $L$ is ample)} \\
			\Rightarrow \left(\frac{d}{d-1} \right)^2 &>& 2 \quad \quad \text{(since $C_1^2 = C_1\cdot C_2 > 0$)}.
		\end{eqnarray*}
		This is a contradiction, since $d\geq 4$.
	\end{proof}

>>>>

	\section{Seshadri constants for certain transversal curve arrangements}\label{main}	
	%\subsection{General results for any surface}

In this section, we will first prove a result computing Seshadri constants for arbitrary transversal arrangements under a nefness condition. We will then study special cases of this theorem.

We first introduce some notation.

	%\begin{definition}
		Let $\mathcal{C}$ be a transversal arrangement of curves on a smooth projective surface $X$.

  For each $1\le i \le d$, let $b_i =  \# \, C_{i} \cap {\rm Sing}(\mathcal{C})$. Since our arrangement is connected, $b_i >0$ for all $i$.
   We define the \emph{base constant} of $\mathcal{C}$ as 
   $$\text{bs}(\mathcal{C}) = \max\{b_1,\dots, b_d\},$$ i.e., 
$\text{bs}(\mathcal{C})$		
  is equal to the maximum number of singular points  of $\mathcal{C}$ that are contained in a single curve $C_{i} \in \mathcal{C}$.
	%\end{definition}

	\begin{question}\label{question}
		Let $\mathcal{C} =\{C_1,C_2, \ldots ,C_d\}$ be a transversal arrangement of curves on a smooth projective surface $X$. Let $L$ be an ample line bundle on $X$. Is it true that
		\begin{equation*}\varepsilon(X, L,{\rm Sing}(\mathcal{C})) =  \inf_{C_i}{\left{\frac{L\cdot C_i}{b_i}\}\right}?
		\end{equation*}
	\end{question}

	For star arrangements, the base constant is exactly equal to $C_1^2(d-1)$. We answer this question positively in Corollary \ref{prop:star} for some arrangement.

The following is our main theorem. It gives the value of Seshadri constant for an arbitrary transversal configuration 
$\mathcal{C} = \{C_1, \dots, C_d\}$
under a nefness condition.

 %We now show that the lower bound in Question \ref{QuestM} can be achieved by certain star configuration of curves on $X.$	So Question \ref{question} has an affirmative answer for these star arrangements.

	%\textcolor{red}{Let us denote $b_i =  \# \, C_{i} \cap {\rm Sing}(\mathcal{C})$. Note that since our arrangement is connected, $b_i >0$.}

	\begin{theorem}
		\label{main-theorem} Let $X$ be a nonsingular projective surface and let $\mathcal{C} = \{C_{1}, \ldots, C_{d}\}$ be a transversal configuration of curves on $X$ with $d \geq 4$. 
		Let $L$ be an ample line bundle on $X$
		such that  $\frac{d\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L-C_i$ is nef for each $i$.   Then 
		%$$\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \inf_{C_i}{\frac{L\cdot C_i}{C_i^2(d-1)}}.$$
		$$\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \inf_{C_i}{\{\frac{L\cdot C_i}{b_i}\}}.$$
	\end{theorem}
	\begin{proof}
	 Let $\mathcal{P} = \{p_{1}, \ldots, p_{s}\}$ be the singular locus of  $\mathcal{C}$. Note that $p_{1}, \ldots, p_{s}$ are all double points of 
	  $\mathcal{C}$. Exactly $b_i$ of these points lie on $C_i$. So the Seshadri ratio computed by $C_i$ is precisely 
	  $ \frac{L\cdot C_i}{b_i}$. %Note that the same ratio is also contributed by each of the curves $C_2,\ldots, C_d$.

Now we will work with configurations satisfying some additional assumptions. More specifically, we will make the following assumption.

	\begin{assumption}\label{star1}
		Let $\mathcal{C}=\{C_1,C_2, \ldots ,C_d\}$, $d\geq 4$ be a (connected) 
  transversal arrangement of curves 
		on a complex smooth projective surface $X$ such that all the curves 
  $C_i$ in $\mathcal{C}$ are linearly equivalent to a
			fixed divisor $A$ on $X$.
 % with the following properties: 
%		\begin{itemize}
%			\item There is no point where all the curves meet, i.e., $t_d=0.$
%			\item  All the curves $C_i$ in $\mathcal{C}$ are linearly equivalent to a
%			fixed divisor $A$ on $X$.
%		\end{itemize}
	\end{assumption}

@@

	For star arrangements, the base constant is exactly equal to $C_1^2(d-1)$. We answer this question positively in Corollary \ref{prop:star} for some arrangement.

We now prove some corollaries for star arrangements where all the curves are linearly equivalent to a fixed divisor. This answers Question \ref{QuestM} in the affirmative.

	\begin{corollary}
		\label{prop:star} Let $X$ be a nonsingular projective surface and let $\mathcal{C} = \{C_{1}, \ldots, C_{d}\}$ be a star configuration of curves on $X$ with $d \geq 4$ such that each $ C_i$ is linearly equivalent to a fixed divisor $A$ on $X$. Let $L$ be an ample line bundle on $X$
		such that  $\frac{dC_1^2}{(d-1)(L\cdot C_1)}L-C_1$ is nef.   Then 
		$$\varepsilon(X, L,{\rm Sing}(\mathcal{C})) = \frac{L\cdot C_1}{C_1^2(d-1)}.$$
	\end{corollary}

 \begin{proof}
 The proof is immediate from Theorem \ref{main-theorem}.    
 \end{proof}
